---
title: "R Coding Exercise"
---
Loading dslabs

```{r}
library(dslabs)
```

Observing gapminder data

```{r}
help(gapminder)

str(gapminder)
summary(gapminder)
class(gapminder)
```

Separating and Observing African countries 

```{r}
africadata<- gapminder[gapminder$continent == "Africa", ]
str(africadata)
summary(africadata)
```

Creating new object for infant mortality and life expectancy 

```{r}
infant_mort_and_Life_Ex <- africadata[, c("infant_mortality", "life_expectancy")]
```


Creating new object for population and life expectancy 

```{r}
pop_and_Life_Ex <- africadata[, c("population", "life_expectancy")]
```


View infant mortality
```{r}
str(infant_mort_and_Life_Ex)
summary(infant_mort_and_Life_Ex)
```


View Popultaion

```{r}
str(pop_and_Life_Ex)
summary(pop_and_Life_Ex)
```



Create plots to view infant mortality and population compared to life expectancy
```{r}
attach(africadata)
plot(infant_mortality, life_expectancy, main="infant mortality and life expectancy", xlab="Infant mortality", ylab= "life_expectancy")


attach(africadata)
plot(infant_mortality, life_expectancy, main="infant mortality and life expectancy", xlab="Infant mortality", ylab= "life_expectancy")

attach(africadata)
plot(log, population, life_expectancy, main="Population and life expectancy", xlab="population", ylab= "life_expectancy")
```



After viewing trends we see there is missing data. Let's take a look at a specfic year, 2000
```{r}
africadata_2000 <-africadata[africadata$year==2000, ]
```



Now lets plot the data again
```{r}
attach(africadata_2000)
plot(infant_mortality, life_expectancy, main="infant mortality and life expectancy", xlab="Infant mortality", ylab= "life_expectancy")

attach(africadata_2000)
plot(log, population, life_expectancy, main="Population and life expectancy", xlab="population", ylab= "life_expectancy")
```


Lets create a linear model from the year 2000 data
```{r}
fit1<-lm(life_expectancy~ infant_mortality, africadata_2000)
summary(fit1)

fit2<-lm(life_expectancy~ population, africadata_2000)
summary(fit2)
```


Infant mortality is a statistically significant predictor of life expectancy, while population was not a statistically significant predictor.


This section was created by Emma Hardin-Parker 

Loading the dataset & packages needed for analysis
```{r}
library(dslabs)
library(tidyverse)
library(ggplot2)

help("mice_weights")
```

Viewing the structure and summary of mice_weights

```{r}
View(mice_weights)
str(mice_weights)
summary(mice_weights)
class(mice_weights)
```

Making various plots to show relationship between bone_density and percent_fat

```{r}
ggplot(mice_weights, aes(x = bone_density, y = percent_fat)) +
  geom_point()
```

From this plot we can see that there is a slight negative relationship between percent_fat and bone_density; as bone density increases, percent_fat tends to decrease. 

I now want to see which litter had higher body weight, stratified by sex. 

```{r}
body_weight_litter <- ggplot(mice_weights, aes(sex, body_weight)) +
                  geom_boxplot(aes(fill = litter))
```

From this plot we can see that the average body_weight in male and female mice are very similar, however, the upper quartile of each boxplot suggests that litter 1 had higher body_weights than litter 2. Female mice in litter 2 had higher body weights on average than litter 1. 

Now, I am going to create a similar plot, however this time I will be seeing the difference in bone_density between male and female mice across both litters. 

```{r}
bone_dens_litter <- ggplot(mice_weights, aes(sex, bone_density)) +
                  geom_boxplot(aes(fill = litter))
```

There are a couple of interesting things to note about this plot. For one, the range is quite large for female mice in litter 1. Conversely from the previous plot, litter 1 had higher bone density across both sexes. 

I now want to see if there is a difference in body weight across litters based on diet. 

```{r}
food_plot <- ggplot(mice_weights, aes(x = diet, y = body_weight)) +
  geom_boxplot(aes(fill = litter))
```

Mice given high fat (hf) diets had consistently higher body weights across both litters and both sexes. 

I now want to perform some simple regressions to see if there are any statistically significant relationships between variables.

```{r}
mice_fit1 <- lm(formula = body_weight ~ bone_density, data = mice_weights)
summary(mice_fit1)
```
The summary of this model suggests that there is a statistically significant relationship between bone_density and body_weight.




